---
title: | 
       | DXE_EMTR Assignment 2 
       | $\scriptsize \color{blue}{\text{"It's}}$ $\scriptsize \color{red}{\text{colorful}}$ $\scriptsize \color{limegreen}{\text{to}}$ $\scriptsize \color{violet}{\text{make}}$ $\scriptsize \color{orange}{\text{it}}$ $\scriptsize \color{SkyBlue}{\text{fun"}} \tiny \text{-- L. Lafférs, 10.11.2022, 14:43, MT205}$
author: "F. Emmer, M. Kutlak, M. Kecskésová"
date: "2022-12-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message=FALSE, echo=FALSE}
library(dagitty)
library(ggdag)
```

## **1 Causal graphical models**
We are interested in quantifying the effect of discrimination on wages. Assume that gender influences the
discrimination but has no direct impact on both occupation and wages. Furthermore, we assume that disc-
rimination influences both occupation and wages. Occupation also have a direct effect on wages. Also,
unobserved ability influences both occupation and wages.  
  
**Task**: Plot the causal graph.  
```{r dags, echo=FALSE}
G1 = dagitty('dag{
W [outcome, pos="2,0"]
D [exposure, pos="1,1"]
A [unobserved, pos="1, -1"]
G [pos="1, 3"]
O [pos="0, 0"]
G -> D
D -> O
D -> W
A -> O
A -> W            
O -> W}')

ggdag(G1) + theme_dag() #plotting the causal graph 
```

**Task**: List all the causal and non-causal paths from discrimination to wages.  
```{r dags1}
paths(G1, 
      from = exposures(G1),
      to = outcomes(G1))
```

-- The causal paths in this case are D -> O -> W and D -> W.  

**Task**: Determine if it is possible to identify the causal effect of discrimination on wages based on observed probability distribution of gender, discrimination, occupation and wages.  
-- There are no testable implications due to both direct and indirect effect of discrimination on wages.

**Task**: Discuss the pros and cons of adjusting for the occupation.  
-- After adjusting for occupation, the path D -> W (the direct effect
 of discrimination on wages) would stay open, however, we would block the indirect path going 
 through occupation, so the resulting effect could be lower. Moreover, occupation
 is a collider variable between discrimination and the unobserved ability,
 so controlling for it opens the path between discrimination, ability and 
 wages and a spurious effect between D and W through A may emerge.  
 
 *Now assume that some occupations are preferred by different genders and that there is a direct link from gender to occupation.*  
 
**Task**: Plot the causal graph.  
```{r dags2, echo=FALSE}
G2 = dagitty('dag{
W [outcome, pos="2,0"]
D [exposure, pos="1,1"]
A [unobserved, pos="1, -1"]
G [pos="1, 3"]
O [pos="0, 0"]
G -> D
G -> O
D -> O
D -> W
A -> O
A -> W            
O -> W}')

ggdag(G2) + theme_dag() #plotting the causal graph 
```

**Task**: List all the causal and non-causal paths from discrimination to wages.  
```{r dags3}
paths(G2, 
      from = exposures(G2),
      to = outcomes(G2))
```

-- The causal paths in this case are D -> O -> W, D -> W and D <- G -> O -> W.  

**Task**: Determine if it is possible to identify the causal effect of discrimination on wages based on observed
probability distribution of gender, discrimination, occupation and wages.  
-- There are again no testable implications due to both direct and indirect effect of discrimination on wages.

**Task**: Discuss the pros and cons of adjusting for the occupation and/or gender.  
**1. adjusting only for occupation:** Same as the previous case, but on top of that a new causal path emerges between gender and wages through ability. The effect of gender on wages possibly also changes due to opening the path between G and W through O and D, as O is also a collider between G and D.  
**2. adjusting only for gender:** Gender influences the indirect path between D and W through O as a confounder influencing D and O. To recover the original indirect path, it is necessary to adjust for gender. 
<!--Adjusting for gender also changes the direct effect of D on W, in case there is an interaction of gender discrimination and other types of discrimination.-->  
**3. adjusting for both occupation and gender:** Adjusting for both gender and occupation closes the indirect path between D and W through O and opens the path between D and W through O and unobserved A. 

*Now furthermore assume that we have a good proxy variable for ability (e.g. composite IQ and EQ score), so that ability is now observed.*

**Task**: Plot the causal graph.  
```{r dags4, echo=FALSE}
G3 = dagitty('dag{
W [outcome, pos="2,0"]
D [exposure, pos="1,1"]
A [observed, pos="1, -1"]
G [pos="1, 3"]
O [pos="0, 0"]
G -> D
G -> O
D -> O
D -> W
A -> O
A -> W            
O -> W}')

ggdag(G3) + theme_dag() #plotting the causal graph 
```

**Task**: List all the causal and non-causal paths from discrimination to wages. 
```{r dags5}
paths(G3, 
      from = exposures(G3),
      to = outcomes(G3))
```

-- The causal paths in this case are D -> O -> W, D -> W and D <- G -> O -> W.  

**Task**: Determine if it is possible to identify the causal effect of discrimination on wages based on observed
probability distribution of gender, discrimination, occupation and wages.  
-- There are again no testable implications due to both direct and indirect effect of discrimination on wages.  

**Task**: Discuss the pros and cons of adjusting for the occupation and/or gender and/or ability.  
**1. occupation only:** -||-  
**2. gender only:** -||-  
**3. ability only:** -||-  
**4. occupation and gender:** -||-  
**5. occupation and ability:** Adjusting for occupation could possibly change the effect of G on D by creating a spurious effect (O is a collider). Adjusting for ability  
**6. gender and ability:** Both direct and indirect effects stay, others are closed  
**7. all three:** For each G we can find the true direct effect of D on W, indirect effect via O 
is blocked. All other paths are closed.  

**Task**: Assume that discrimination variable is binary and propose an estimator for (total) average treatment
effect of discrimination on wages.  
-- Given that by controlling for occupation and ability, we are able to recover both the direct and indirect effect of discrimination on wages, we propose a linear model with OLS estimator, while controlling for occupation and ability. Logit?????? 

**Task**: Complement your analysis of $G_3$ with a small simulation study. Convince the reader that the estimator
that you proposed is able to recover the true (total) average treatment effect of discrimination on
wages. Make any reasonable simplifications in order to illustrate your point.  

```{r estimator}
set.seed(666)
gender <- rbinom(1000, 1, 0.5)
discrimination <- gender + rnorm(1000, 0, 0.2)
discrimination <- ifelse(discrimination >= 0.2, 1, 0)
summary(lm(discrimination ~ gender))

ability <- rnorm(1000)
occupation <- 1.5*gender + discrimination + 2*ability + rnorm(1000, 0, 0.2)
wages <- 3*ability + 2*discrimination + 6*occupation + rnorm(1000, 0, 0.2)

summary(lm(wages ~ discrimination + gender + ability))
```


<center> $\huge \text{"This is depressing, but that's the situation."}$ </center>
  
[$\text{-- L. Lafférs, 10.11.2022, 16:32, MT205}$]{style="float:right"}  

## **2 Selection on Observables**  
Consider the following hypothetical situation. We have a population of 10 individuals and we happen to know
both their contrafactual earnings: if they go to job training programme $(D = 1)$ they will receive $Y(1)$. Or if
they do not participate in the job training programme $(D = 0)$ they will get $Y(0)$. We also have information
on the different background they have: some individuals come from cities $(X = 1)$ while others from rural
areas $(X = 0)$.


<center> $\huge \text{"You should be laughing by now,}$
         $\huge \text{but I understand that you're tired."}$ </center>
  
[$\text{-- L. Lafférs, 10.11.2022, 17:15, MT205}$]{style="float:right"}  